### 2020-06-14 Thesis/games.tex
Update references; one typo; fix broken links.

To introduce our material, we will go back over some old ground, namely the fully abstract game semantics for Idealized Algol developed by Abramsky and McCusker in \cite{SamsonGuyIAActive}.  
In keeping with the spirit of this thesis, we will aim to use category theoretic methods, and so our proofs of soundness and adequacy will depart from those given by Abramsky and McCusker, and will instead involve coalgebraic ideas developed by Laird in \cite{laird02} and \cite{LairdCofCommCom}.

>>>>

To introduce our material, we will go back over some old ground, namely the fully abstract game semantics for Idealized Algol developed by Abramsky and McCusker in \cite{SamsonGuyIAActive}.  
In keeping with the spirit of this thesis, we will aim to use category theoretic methods, and so our proofs of soundness and adequacy will depart from those given by Abramsky and McCusker, and will instead involve coalgebraic ideas developed by Laird in \cite{laird02} and \cite{FunctionalProgramsAsCoroutines}.

### 2020-06-14 Thesis/intro.tex
Update references; one typo; fix broken links.

The last problem is to do with definability.
First note that there are two different reasons why an element of a denotational model (e.g., a strategy) might not be definable in a particular language.  
One reason is structural: for example, a stateless language such as PCF cannot define a strategy whose behaviour depends on its entire history, since that would indicate stateful behaviour.  
The other reason is purely computational.  
Given a non-computable function $f\from \bN \to \bN$ (for example, the function that returns $0$ if its argument is source code for a terminating program and $1$ otherwise), we have a perfectly well-behaved history free strategy that represents $f$, but $f$ is nevertheless still not definable in most programming languages.  
Historically, the study of Full Abstraction has not been too concerned with non-definable elements of the second kind: the reason is that they do not usually play a role in determining the intrinsic equivalence relation.
However, in the presence of countable nondeterminism, there exist definable terms that can be distinguished in the model that cannot be distinguished by definable elements, but can be distinguished by non-computable functions.
An example of such a function is found in Proposition \ref{PropNotFullyAbstract}; the general idea is due to Kleene and can be found in \cite{KleeneTree}.

>>>>

The last problem is to do with definability.
First note that there are two different reasons why an element of a denotational model (e.g., a strategy) might not be definable in a particular language.  
One reason is structural: for example, a stateless language such as PCF cannot define a strategy whose behaviour depends on its entire history, since that would indicate stateful behaviour.  
The other reason is purely computational.  
Given a non-computable function $f\from \bN \to \bN$ (for example, the function that returns $0$ if its argument is source code for a terminating program and $1$ otherwise), we have a perfectly well-behaved history free strategy that represents $f$, but $f$ is nevertheless still not definable in most programming languages.  
Historically, the study of Full Abstraction has not been too concerned with non-definable elements of the second kind: the reason is that they do not usually play a role in determining the intrinsic equivalence relation.
However, in the presence of countable nondeterminism, there exist definable terms that can be distinguished in the model that cannot be distinguished by definable elements, but can be distinguished by non-computable functions.
An example of such a function is found in Section \ref{SecNonComputable}; the general idea is due to Kleene and can be found in \cite{KleeneTree}.

### 2020-06-14 Thesis/pmonads.tex
Update references; one typo; fix broken links.

  (ii): The colimit is given by the set
  \[
    c = \sum_{d \in \D} J(d) / \sim\,,
    \]
  where $\sim$ is the equivalence relation on pairs $(d, \eta_d)$ (for $d\in \D$ and $\eta_d\in J(d)$), generated by
  \[
    \left(\exists y\from \YY, f\from d \to y,g\from e \to y \esuchthat J(f)(\eta_d) = J(g)(\eta_{e})\right) \mathrel{\Rightarrow} (d,\eta_d) \sim (e,\eta_e)\,.
    \]
  The proofs that these are indeed a limit and a colimit for the corresponding diagrams are very similar to the proofs for ordinary small limits and colimits in $\Set$.  
  However, the constructions we have made above rely on more than the fact that $\Set$ is complete and cocomplete.  
  We also need to be able to perform universal and existential quantification over large collections of objects; this corresponds to the category-theoretic fact that if we have any (possibly large) collection of monomorphisms
  \[
    A_i \to B
    \]
  in $\Set$ (for fixed $B$), then we can form their intersection (i.e., a category-theoretic pullback for the collection), and similarly that we can form the pushout of any (possibly large) collection of epimorphisms out of a set.
  See \ref{AllGoodCategories} for full details.
\end{proof}

@@

Note that these results is not the sharpest possible: there may well be more complicated zigzags mediating between the morphisms.
Nevertheless, they are sufficient for our purposes.

@@

\begin{example}
  An alternative way of dealing with probability that does satisfy the condition for Cartesianness of the resulting category.
  If $(\Omega,\F,\bP)$ is a fixed probability space, then a \emph{discrete random variable taking values in a set $A$} is a measurable function
  \[
    V\from (\Omega,\F) \to (A,\powerset A)\,.
    \]
  Given $X \subset A$, we define
  \[
    \bP(V \in X) = \bP(V\inv(X))\,.
    \]
  This then gives us a discrete probability measure on $X$.

>>>>

  (ii): The colimit is given by the set
  \[
    c = \sum_{d \in \D} J(d) / \sim\,,
    \]
  where $\sim$ is the equivalence relation on pairs $(d, \eta_d)$ (for $d\in \D$ and $\eta_d\in J(d)$), generated by
  \[
    \left(\exists y\from \YY, f\from d \to y,g\from e \to y \esuchthat J(f)(\eta_d) = J(g)(\eta_{e})\right) \mathrel{\Rightarrow} (d,\eta_d) \sim (e,\eta_e)\,.
    \]
  The proofs that these are indeed a limit and a colimit for the corresponding diagrams are very similar to the proofs for ordinary small limits and colimits in $\Set$.  
  However, the constructions we have made above rely on more than the fact that $\Set$ is complete and cocomplete.  
  We also need to be able to perform universal and existential quantification over large collections of objects; this corresponds to the category-theoretic fact that if we have any (possibly large) collection of monomorphisms
  \[
    A_i \to B
    \]
  in $\Set$ (for fixed $B$), then we can form their intersection (i.e., a category-theoretic pullback for the collection), and similarly that we can form the pushout of any (possibly large) collection of epimorphisms out of a set.
  See \cite{AllGoodCategories} for full details.
\end{proof}

@@

Note that these results are not the sharpest possible: there may well be more complicated zigzags mediating between the morphisms.
Nevertheless, they are sufficient for our purposes.

@@

\begin{example}
  \label{ExRandomVariable}
  An alternative way of dealing with probability that does satisfy the condition for Cartesianness of the resulting category.
  If $(\Omega,\F,\bP)$ is a fixed probability space, then a \emph{discrete random variable taking values in a set $A$} is a measurable function
  \[
    V\from (\Omega,\F) \to (A,\powerset A)\,.
    \]
  Given $X \subset A$, we define
  \[
    \bP(V \in X) = \bP(V\inv(X))\,.
    \]
  This then gives us a discrete probability measure on $X$.
